---
title: "Cleaning Location Data of Wolves Fishing in Northern Minnesota"
author: "Dani Freund"
format: 
  html:
    theme: lux
    embed-resources: TRUE
echo: TRUE
warning: FALSE
error: FALSE
editor: visual
---

## Analysis Notes

-   don't have full season location data for some wolves

-   O0C doesn't have an after fishing period

-   V067 does not have any locations

-   V034 does not have locations for 2019

### Packages Needed

```{r}
library(readr) # for read_csv
library(tidyr) # for data organization
library(dplyr) # for data organization
library(lubridate) # for dates
library(rgdal) # for shapefiles
library(ggplot2) # for plotting
library(kableExtra) # for making tables
```

#### Loading Data

```{r output = FALSE, warning = FALSE}
locations = read.csv(file = "/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/all_wolves_locs.csv")

fishing.periods = read.csv(file = "/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/summary_fishing_periods_final.csv")

buff.Ash.2 <-readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/ash_river_20m_buffer/AshRiverBuffer20Buffer.shp")
buff.Irwin.2 <-readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/irwin_creek_20m_buffer/IrwinCreek20Buffer.shp")
buff.Daley.2 <-readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/daley_creek_20m_buffer/daley.Creek.Buffer.20.shp")
buff.Black.2 <- readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/black_duck_river_20m_buffer/black.Duck.River.Buffer.20.shp")

buff.Ash.5 <-readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/ash_river_500m_buffer/AshRiver500Buffer.shp")
buff.Irwin.5 <-readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/irwin_creek_500m_buffer/irwin.Creek.Buffer.500m.shp")
buff.Daley.5 <-readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/daley_creek_500m_buffer/daley.Creek.Buffer.500.shp")
buff.Black.5 <- readOGR(dsn="/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/waterway_buffers/black_duck_river_500m_buffer/black.Duck.River.Buffer.500.shp")
```

### Summarizing Entire Data Set

```{r}
locations$date<-as.Date(locations$date) # need this to do calculations with dates

# getting max/min dates for entire data set for each wolf 
#use the tidyverse functions (specifically functions from dplyr package) to summarize data in a variety of ways
#note: the group_by() functions groups the data by wolf and then by date for each wolf.
locations.all.data<-locations %>% group_by(wolf) %>% 
    summarize(numberLocations = n(), #counts up the number of rows (i.e., locations) #"numberLocations" is the column name that the count data goes in
              start = min(date), #selects the earliest date
              end = max(date), #selects the latest date
              length = max(date)-min(date)) #time between the two

locations.all.data %>% kbl() %>% kable_material_dark(c("striped", "hover"))
```

### Cleaning Data

```{r}
O0C.during.min = '2021-05-26' %>% as.Date()
O0C.during.max = "2021-07-07" %>% as.Date()

V046.during.min = "2017-04-21" %>% as.Date()
V046.during.max = "2017-05-16" %>% as.Date()

V060.during.min = "2018-05-12" %>% as.Date()
V060.during.max = "2018-05-18" %>% as.Date()

V062.during.min = "2018-05-09" %>% as.Date()
V062.during.max = "2018-05-25" %>% as.Date()

V071.during.min = "2021-04-24" %>% as.Date()
V071.during.max = "2021-05-17" %>% as.Date()
  
V077.during.min = "2020-05-19" %>% as.Date()
V077.during.max = "2020-05-29" %>% as.Date()
  
V089.during.min = "2020-05-20" %>% as.Date() # changed to just using camera collar data
V089.during.max = "2020-05-28" %>% as.Date()
  
V094.during.min = "2021-05-02" %>% as.Date()
V094.during.max = "2021-05-09" %>% as.Date()

V034.during.min = "2017-04-17" %>% as.Date()
V034.during.max = "2017-05-17" %>% as.Date()

# finding before and after periods
before.period = function(min,max){
          length=max-min
          before=min-length
}
after.period = function(min,max){
        length=max-min
        after=max+length
}

O0C.before=before.period(O0C.during.min,O0C.during.max)
O0C.after=after.period(O0C.during.min,O0C.during.max)

V046.before=before.period(V046.during.min,V046.during.max)
V046.after=after.period(V046.during.min,V046.during.max)

V060.before=before.period(V060.during.min,V060.during.max)
V060.after=after.period(V060.during.min,V060.during.max)

V062.before=before.period(V062.during.min,V062.during.max)
V062.after=after.period(V062.during.min,V062.during.max)

V071.before=before.period(V071.during.min,V071.during.max)
V071.after=after.period(V071.during.min,V071.during.max)

V077.before=before.period(V077.during.min,V077.during.max)
V077.after=after.period(V077.during.min,V077.during.max)

V089.before=before.period(V089.during.min,V089.during.max)
V089.after=after.period(V089.during.min,V089.during.max)

V094.before=before.period(V094.during.min,V094.during.max)
V094.after=after.period(V094.during.min,V094.during.max)

V034.before=before.period(V034.during.min,V034.during.max)
V034.after=after.period(V034.during.min,V034.during.max)

# filtering out periods
# for functions to filter wolfID properly, need to but ' ' in the call to function, not when writing the function
filter.period = function(data,wolfID,min,max){
  wolf=filter(data, wolf==wolfID & date >= min & date <= max)}

O0C.during.period = filter.period(locations, 'O0C', O0C.during.min, O0C.during.max) %>% mutate(period='during')
O0C.after.period = filter.period(locations, 'O0C', O0C.during.max, O0C.after) %>% mutate(period='after')
# O0C doesn't have an after period

V046.before.period = filter.period(locations, 'V046', V046.before, V046.during.min) %>% mutate(period='before')
V046.during.period = filter.period(locations, 'V046', V046.during.min, V046.during.max) %>% mutate(period='during')
V046.after.period = filter.period(locations, 'V046', V046.during.max, V046.after) %>% mutate(period='after')

V060.before.period = filter.period(locations, 'V060', V060.before, V060.during.min) %>% mutate(period='before')
V060.during.period = filter.period(locations, 'V060', V060.during.min, V060.during.max) %>% mutate(period='during')
V060.after.period = filter.period(locations, 'V060', V060.during.max, V060.after) %>% mutate(period='after')

V062.before.period = filter.period(locations, 'V062', V062.before, V062.during.min) %>% mutate(period='before')
V062.during.period = filter.period(locations, 'V062', V062.during.min, V062.during.max) %>% mutate(period='during')
V062.after.period = filter.period(locations, 'V062', V062.during.max, V062.after) %>% mutate(period='after')

V071.before.period = filter.period(locations, 'V071', V071.before, V071.during.min) %>% mutate(period='before')
V071.during.period = filter.period(locations, 'V071', V071.during.min, V071.during.max) %>% mutate(period='during')
V071.after.period = filter.period(locations, 'V071', V071.during.max, V071.after) %>% mutate(period='after')

V077.before.period = filter.period(locations, 'V077', V077.before, V077.during.min) %>% mutate(period='before')
V077.during.period = filter.period(locations, 'V077', V077.during.min, V077.during.max) %>% mutate(period='during')
V077.after.period = filter.period(locations, 'V077', V077.during.max, V077.after) %>% mutate(period='after')

V089.before.period = filter.period(locations, 'V089', V089.before, V089.during.min) %>% mutate(period='before')
V089.during.period = filter.period(locations, 'V089', V089.during.min, V089.during.max) %>% mutate(period='during')
V089.after.period = filter.period(locations, 'V089', V089.during.max, V089.after) %>% mutate(period='after')

V094.before.period = filter.period(locations, 'V094', V094.before, V094.during.min) %>% mutate(period='before')
V094.during.period = filter.period(locations, 'V094', V094.during.min, V094.during.max) %>% mutate(period='during')
V094.after.period = filter.period(locations, 'V094', V094.during.max, V094.after) %>% mutate(period='after')

V034.before.period = filter.period(locations, 'V034', V034.before, V034.during.min) %>% mutate(period='before')
V034.during.period = filter.period(locations, 'V034', V034.during.min, V034.during.max) %>% mutate(period='during')
V034.after.period = filter.period(locations, 'V034', V034.during.max, V034.after) %>% mutate(period='after')

fishing.periods.summary = rbind(O0C.during.period, O0C.after.period,
                                V046.before.period, V046.during.period, V046.after.period,
                                V060.before.period, V060.during.period, V060.after.period,
                                V062.before.period, V062.during.period, V062.after.period,
                                V071.before.period, V071.during.period, V071.after.period,
                                V077.before.period, V077.during.period, V077.after.period,
                                V089.before.period, V089.during.period, V089.after.period,
                                V094.before.period, V094.during.period, V094.after.period,
                                V034.before.period, V034.during.period, V034.after.period) %>% 
  mutate(river ='NA', buffer='none')

write.csv(fishing.periods.summary,"/Users/danifreund/Desktop/fishing_data_code_for_publication/ethology_of_wolves_freshwater_fishing/data/all.locs.fishing.period.no.buffers.csv")

############### FILTERING OUT WOLF LOCATIONS BY EACH WOLF'S RESPECTIVE BUFFER

filter.by.buff=function(data,wolfID,buffer){
  filter.locs=filter(data, wolf==wolfID)
  coordinates(filter.locs) <- ~longitude.Deg + latitude.Deg #'WGS 84 LONGITUDE is before LATITUDE!!!!
  proj4string(filter.locs) <- CRS("+init=epsg:4326")
  filter.locs<-spTransform(filter.locs, CRS(proj4string(buffer))) # converts locations to UTM
  buff.locs<-filter.locs[buffer,] 
  plot(buff.locs, col='red') 
  plot(buffer, add=T) 
  data.frame(buff.locs)}

O0C.ash.5 = filter.by.buff(fishing.periods.summary,'O0C', buff.Ash.5) %>% mutate(river = "ash", buffer= "500")
O0C.ash.2 = filter.by.buff(fishing.periods.summary,'O0C', buff.Ash.2) %>% mutate(river = "ash", buffer= "20")
O0C.black.5 = filter.by.buff(fishing.periods.summary,'O0C', buff.Black.5) %>% mutate(river = "black", buffer= "500")
O0C.black.2 = filter.by.buff(fishing.periods.summary,'O0C', buff.Black.2) %>% mutate(river = "black", buffer= "20")

V046.irwin.5 = filter.by.buff(fishing.periods.summary,'V046', buff.Irwin.5) %>% mutate(river = "irwin", buffer= "500")
V046.irwin.2 = filter.by.buff(fishing.periods.summary,'V046', buff.Irwin.2) %>% mutate(river = "irwin", buffer= "20")

V060.irwin.5 = filter.by.buff(fishing.periods.summary,'V060', buff.Irwin.5) %>% mutate(river = "irwin", buffer= "500")
V060.irwin.2 = filter.by.buff(fishing.periods.summary,'V060', buff.Irwin.2) %>% mutate(river = "irwin", buffer= "20")

V062.irwin.5 = filter.by.buff(fishing.periods.summary,'V062', buff.Irwin.5) %>% mutate(river = "irwin", buffer= "500")
V062.irwin.2 = filter.by.buff(fishing.periods.summary,'V062', buff.Irwin.2) %>% mutate(river = "irwin", buffer= "20")

V071.daley.5 = filter.by.buff(fishing.periods.summary,'V071', buff.Daley.5) %>% mutate(river = "daley", buffer= "500")
V071.daley.2 = filter.by.buff(fishing.periods.summary,'V071', buff.Daley.2) %>% mutate(river = "daley", buffer= "20")

V077.ash.5 = filter.by.buff(fishing.periods.summary,'V077', buff.Ash.5) %>% mutate(river = "ash", buffer= "500")
V077.ash.2 = filter.by.buff(fishing.periods.summary,'V077', buff.Ash.2) %>% mutate(river = "ash", buffer= "20")

V089.ash.5 = filter.by.buff(fishing.periods.summary,'V089', buff.Ash.5) %>% mutate(river = "ash", buffer= "500")
V089.ash.2 = filter.by.buff(fishing.periods.summary,'V089', buff.Ash.2) %>% mutate(river = "ash", buffer= "20")

V094.irwin.5 = filter.by.buff(fishing.periods.summary,'V094', buff.Irwin.5) %>% mutate(river = "irwin", buffer= "500")
V094.irwin.2 = filter.by.buff(fishing.periods.summary,'V094', buff.Irwin.2) %>% mutate(river = "irwin", buffer= "20")

# don't have to specify year here because it will capture both 2017 and 2019
V034.irwin.5 = filter.by.buff(fishing.periods.summary,'V034', buff.Irwin.5) %>% mutate(river = "irwin", buffer= "500")
V034.irwin.2 = filter.by.buff(fishing.periods.summary,'V034', buff.Irwin.2) %>% mutate(river = "irwin", buffer= "20")

summary.locs.buffered = rbind(O0C.ash.5, O0C.ash.2, O0C.black.5, O0C.black.2, 
                              V046.irwin.5, V046.irwin.2, 
                              V060.irwin.5, V060.irwin.2,
                              V062.irwin.5, V062.irwin.2, 
                              V071.daley.5, V071.daley.2, 
                              V077.ash.5, V077.ash.2, 
                              V089.ash.5, V089.ash.2, 
                              V094.irwin.5, V094.irwin.2,
                              V034.irwin.5, V034.irwin.2)

# checking to make sure it worked
summary.locs.buffered %>% subset(period == "during") %>% group_by(wolf,river,buffer) %>% summarize(start=min(date),end=max(date)) %>% kbl() %>% kable_material_dark(c("striped", "hover"))

# saving out csv
write.csv(summary.locs.buffered, "/Users/danifreund/Desktop/fishing_data_code_for_publication/wolves_fishing/data/locations_by_fishing_period.csv")
```

V060 had a 12hour collar, so the location dates do not match up with the camera trap dates that were used to define their fishing period. However, all location dates in this data set are within the fishing period dates from the camera trap dates used to define the fishing period.
